Relationship between solar wind, $D_{st}$, and plasmasphere trough mass density on one-hour time scales

### 2016-03-04 paper.tex
Updating tableparse to also do nonlinear perltable
misc

\section{Results}

\subsection{$D_{st}$ and $\rho_{eq}$ Events}

@@

@@

In the upper panel of Figure \ref{DailyAverages}, $D_{st}$ events are shown.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is $\sim 2$ amu/cm$^3$ instead of the value of $\sim 10$ amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for the non-\req\ medians, the number of measurements used in computing the medians is typically equal to the number of events which is much larger than the number of \req\ values as we use all available measurements instead of restricting to only values where a non--fill $\rho_{eq}$ value existed. \vinote{Alternate: Note that for $D_{st}$ events, we computed median values using all available data, even when \req\ availability was sparse.}

In the lower panel of Figure \ref{DailyAverages} it is shown that all trends seen in the analysis of 1989-1991 also hold for the entire run of GOES 6 over 1983-1991, including the magnitude of change in $\rho_{eq}$ across event onset.

@@

This method finds 669 periods of $D_{st}$ crossing beneath the selected threshold between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours \vinote{Make sure I defined "duration"}. Figure \ref{Storms} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{Storms}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40$~nT. The dashed red lines indicate plus and minus one standard deviation of values from all events that were used to compute the median. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack.

Figure \ref{Storms} shows a definite spike in the $Z$ component of the magnetic field, as well as the defined drop in $D_{st}$, but no obvious change in mass density at an hourly timescale. This points to an issue with only looking at long-timescale trends between density and $D_{st}$, and allows for the possibility that other factors are influencing the long term correlation since there's no obvious connection on a short timescale. One possibility is that, as suggested in \cite{Takahashi2010}, $F_{10.7}$ plays a significant role in driving long term density values which biases the long term correlation of density and $D_{st}$.

@@

In the upper panel of \ref{DailyAverages}, $D_{st}$ events are shown.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events corrspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is on the order of $\sim2$~amu/cm$^3$ instead of $\sim10$~amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for $D_{st}$, the number of measurements used in computing the medians is much larger as we use all available $D_{st}$ measurements instead of restricting to only values where a non-fill $\rho_{eq}$ value existed.

In the lower panel of \ref{DailyAverages}, $\rho_{eq}$ events are shown.  Here we see that $\rho_{eq}$ enhancements do not clearly correspond to a signature in $D_{st}$, but that there is a trend for such enhancments to occur when $B_{z}$ in the IMF is positive.  This is consistent with \cite{Takahashi2010} who note that very large increases in $\rho_{eq}$ often corresponded to movement of the plasmasphere to beyond geosynchronous distances.

A total of 668 events were found in the interval of May 1983 through August 1991 with an average duration of 9 hours and a median duration of 3 hours. Figure \ref{Storms} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{Storms}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-50$~nT. The dashed red lines indicate plus and minus one standard deviation of values from all events that went into the average. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack. A subset of longer duration events will be looked at later.

@@

\subsection{Change in $\rho_{eq}$}
Instead of looking for events based on a threshold of $D_{st}$, we instead looked for a certain amount of change in $\rho_{eq}$ as the basis for event onset. Both raw change and percent change were done in case of bias towards high or low $\rho_{eq}$ periods, respectively, but neither showed any correlation to a change in $D_{st}$.

@@

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-50-tak.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-day.eps}
\caption{$D_{st}$ events from GOES 6. Top: Events in the interval 1989-1991; compare to \cite{Takahashi2010} Fig.~11. Bottom: Events in the interval 1983-1991.}
\label{DailyAverages}
\end{figure}

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dd12.eps}
\caption{(a) Average of solar wind and near-Earth measurements around the time $D_{st}$ crossed below $-50$~nT onset. (b) Same as (a) except for constraint that $D_{st}$ stayed below $-50$~nT for at least 12~hours.}
\label{Storms}
\end{figure}

@@

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-d80.eps}
\rule[1ex]{5cm}{1pt}
\caption{(a) Same as Figure \ref{Storms}(a) except with constraint that $D_{st}$ crossed below $-80$~nT }
\label{Dspec}
\end{figure}

>>>>

\section{Epoch Average Results}

@@

\subsection{$D_{st}$ Events}

@@

In the upper panel of Figure \ref{DailyAveragedDstEvents}, $D_{st}$ events are shown.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is $\sim 2$ amu/cm$^3$ instead of the value of $\sim 10$ amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for the non-\req\ medians, the number of measurements used in computing the medians is typically equal to the number of events which is much larger than the number of \req\ values as we use all available measurements instead of restricting to only values where a non--fill $\rho_{eq}$ value existed. \vinote{Alternate: Note that for $D_{st}$ events, we computed median values using all available data, even when \req\ availability was sparse.}

In the lower panel of Figure \ref{DailyAveragedDstEvents} it is shown that all trends seen in the analysis of 1989-1991 also hold for the entire run of GOES 6 over 1983-1991, including the magnitude of change in $\rho_{eq}$ across event onset.

@@

This method finds 669 periods of $D_{st}$ crossing beneath the selected threshold between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours \vinote{Make sure I defined "duration"}. Figure \ref{HourlyAveragedDstEvents} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{HourlyAveragedDstEvents}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40$~nT. The dashed red lines indicate plus and minus one standard deviation of values from all events that were used to compute the median. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack.

Figure \ref{HourlyAveragedDstEvents} shows a definite spike in the $Z$ component of the magnetic field, as well as the defined drop in $D_{st}$, but no obvious change in mass density at an hourly timescale. This points to an issue with only looking at long-timescale trends between density and $D_{st}$, and allows for the possibility that other factors are influencing the long term correlation since there's no obvious connection on a short timescale. One possibility is that, as suggested in \cite{Takahashi2010}, $F_{10.7}$ plays a significant role in driving long term density values which biases the long term correlation of density and $D_{st}$.

@@

In the upper panel of \ref{DailyAveragedDstEvents}, $D_{st}$ events are shown.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events corrspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is on the order of $\sim2$~amu/cm$^3$ instead of $\sim10$~amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for $D_{st}$, the number of measurements used in computing the medians is much larger as we use all available $D_{st}$ measurements instead of restricting to only values where a non-fill $\rho_{eq}$ value existed.

In the lower panel of \ref{DailyAveragedDstEvents}, $\rho_{eq}$ events are shown.  Here we see that $\rho_{eq}$ enhancements do not clearly correspond to a signature in $D_{st}$, but that there is a trend for such enhancments to occur when $B_{z}$ in the IMF is positive.  This is consistent with \cite{Takahashi2010} who note that very large increases in $\rho_{eq}$ often corresponded to movement of the plasmasphere to beyond geosynchronous distances.

A total of 668 events were found in the interval of May 1983 through August 1991 with an average duration of 9 hours and a median duration of 3 hours. Figure \ref{HourlyAveragedDstEvents} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{HourlyAveragedDstEvents}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-50$~nT. The dashed red lines indicate plus and minus one standard deviation of values from all events that went into the average. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack. A subset of longer duration events will be looked at later.

@@

\subsection{$\rho_{eq}$ Events}

Instead of looking for events based on a threshold of $D_{st}$, we instead looked for a certain amount of change in $\rho_{eq}$ as the basis for event onset. Both raw change and percent change were done in case of bias towards high or low $\rho_{eq}$ periods, respectively, but neither showed any correlation to a change in $D_{st}$.

An alterntive type of event can be defined that corresponds to large increases in $\rho\ $.  Figure~\ref{HourlyAveragedRhoEvents} shows the epoch time series for events in which an increase of $\req\ $ above $20$~amu/cm$^2$ defines the start of the event.  Here we see the expected result that such events tend to be associated with positive $B_z^{IMF}$.  To determine if the sign of $B_z^{IMF}$ influences the $\rho\ $ epoch time series, we separated the events by the average of $B_z^{IMF}$ in a time window of $4$~hours centered on the event start time.  The epoch curves for $B_z^{IMF}\geq=0$ (257) events and $B_z^{IMF}\ge=0$ (139) have no statistically signifiant difference.

@@

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-50-tak.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-day.eps}
\caption{$D_{st}$ events from GOES 6 using daily medians. Top: Events in the interval 1989-1991; compare to \cite{Takahashi2010} Fig.~11. Bottom: Events in the interval 1983-1991.}
\label{DailyAveragedDstEvents}
\end{figure}

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dd12.eps}
\caption{$D_{st}$ events from GOES 6 using hourly medians. Top: Events in the interval 1983-1991; compare to bottom panel {DailyAveragedDstEvents}. Bottom: Same as Top except for constraint that $D_{st}$ stayed below $-50$~nT for at least 12~hours.}
\label{HourlyAveragedDstEvents}
\end{figure}

@@

\begin{figure}[htp!]
	\centering
	\includegraphics[scale=0.40]{paperfigures/stormavs-mass-gt20-GOES6.eps}
\caption{$\req\ > 20$~ amu/cm^3 events from GOES 6 using hourly medians. As described in the text, there is no statistically significant dependence of the $\req\ $ plot on $B_z$ in the time interval of $\pm 2$~hours around the event.}
\label{HourlyAveragedRhoEvents}
\end{figure}

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-d80.eps}
\rule[1ex]{5cm}{1pt}
\caption{(a) Same as Figure \ref{HourlyAveragedDstEvents}(a) except with constraint that $D_{st}$ crossed below $-80$~nT }
\label{Dspec}
\end{figure}
